The Impact of Malaria Episodes and Treatment Regimens on Adverse Pregnancy Outcomes in Ugandan Women

### 2025-03-30 products/manuscript/Manuscript.qmd
Project1
Change

>>>>

*Table 2: Association Between Malaria Episodes and Risk of Preterm Birth*
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Create cleaned table for Table 2
table_preterm_pars_clean <- tribble(
  ~Variable, ~`Adjusted Odds Ratio (aOR)`, ~`95% CI Lower`, ~`95% CI Upper`, ~`p-value`,
  "Total malaria episodes during pregnancy", 1.04, 0.95, 1.14, 0.401,
  "Age at enrollment (years)", 0.99, 0.96, 1.03, 0.697,
  "Gravidity", 0.91, 0.82, 1.02, 0.106,
  "Parity", 1.17, 1.05, 1.30, 0.005,
  "Hypertension", 0.89, 0.65, 1.22, 0.463,
  "Diabetes mellitus", 1.22, 0.88, 1.69, 0.237
)

# Display table with consistent styling
kable(table_preterm_pars_clean,
      caption = "",
      col.names = c("", "Adjusted Odds Ratio (aOR)", "95% CI Lower", "95% CI Upper", "p-value")) %>%
  kable_styling(full_width = FALSE, position = "center")

*Table 3: Association Between Malaria Episodes and Risk of Stillbirth*
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
library(logistf)

# Run parsimonious stillbirth model
model_stillbirth_pars <- logistf(
  stillbirth_bin ~ total_malaria_episodes_during_pregnancy +
    age_at_enrollment_years + gravidity + parity +
    hypertension + diabetes_mellitus,
  data = promo_data_clean
)

# Extract and format results
table_stillbirth_pars <- data.frame(
  Variable = names(model_stillbirth_pars$coefficients),
  `Adjusted Odds Ratio (aOR)` = round(exp(model_stillbirth_pars$coefficients), 2),
  `95% CI Lower` = round(exp(model_stillbirth_pars$ci.lower), 2),
  `95% CI Upper` = round(exp(model_stillbirth_pars$ci.upper), 2),
  `p-value` = signif(model_stillbirth_pars$prob, 3)
)

# Remove the intercept row
table_stillbirth_pars <- table_stillbirth_pars[table_stillbirth_pars$Variable != "(Intercept)", ]

# Map raw variable names to clean, reader‐friendly labels
variable_labels <- c(
  "total_malaria_episodes_during_pregnancy" = "Total malaria episodes during pregnancy",
  "age_at_enrollment_years" = "Age at enrollment (years)",
  "gravidity" = "Gravidity",
  "parity" = "Parity",
  "hypertensionYes" = "Hypertension",
  "diabetes_mellitusYes" = "Diabetes mellitus"
)

# Replace raw names with neat labels; replace any NA (i.e., unmapped) with blank
table_stillbirth_pars$Variable <- variable_labels[table_stillbirth_pars$Variable]
table_stillbirth_pars$Variable[is.na(table_stillbirth_pars$Variable)] <- ""

# Remove any row names so only the column appears in the final table
rownames(table_stillbirth_pars) <- NULL

# Display table with consistent styling and without row names
kable(table_stillbirth_pars,
      row.names = FALSE,
      caption = "*Table 3: Stillbirth Model*",
      col.names = c("Exposure and Adjustment Variables", "Adjusted Odds Ratio (aOR)", "95% CI Lower", "95% CI Upper", "p-value")) %>%
  kable_styling(full_width = FALSE, position = "center")

*Association between Malaria Episodes and Adverse Birth Outcomes*
In adjusted logistic regression analysis (Table 4a), the number of malaria episodes during pregnancy was not significantly associated with preterm birth (aOR: 1.04; 95% CI: 0.95–1.14; p = 0.401). Similarly, the penalized Firth logistic regression model showed no significant association between malaria episodes and stillbirth (aOR: 0.96; 95% CI: 0.83–1.10; p = 0.559) (Table 4b). No significant associations were observed with age at enrollment, gravidity, or hypertension for either outcome.

*Effect of Treatment Arm on Malaria Episode Incidence*
A Poisson regression model assessing malaria episode incidence by treatment arm found no significant difference between the DP and SP treatment groups (IRR: 1.02; 95% CI: 0.92–1.12; p = 0.756) after adjusting for age, gravidity, and education level (Table 4c).

A summary of these primary outcome analyses is presented in Table 4.
